Fourmile Calibrations

### 2023-01-30 calibration_output.Rmd
added: map, boulder creek n75th gage, maxQ prior to Apr1, compairson of niwot & hru7 precip, comparison of niwot % hru7 ET

library(tidyverse)
library(plotly)
library(lubridate)
library(knitr)
library(kableExtra)
library(dataRetrieval)
library(hydroGOF)

@@

@@

@@

q_ensemble <- ggplot(ensem_pivot, aes(date, q_m3s, color = cal_run)) +
  geom_line() +
  scale_color_manual(name="cal_run", values=c('obs'='black',
                                            "grid_ensemble"="maroon", 
                                            "statn_ensemble"="darkturquoise", 
                                            "grid_all"="purple1",
                                            "statn_all"='springgreen4', 
                                            "all_ensemble" = "orange"))

ggplotly(q_ensemble)

@@

plotqensem_cy <- function(calyear) {
  {ensem_pivot %>%
    filter(year == calyear) %>%
    ggplot(aes(x = date,
               y = q_m3s,
               color = cal_run)) +
    geom_line() +
    scale_color_manual(name="cal_run", values=c('obs' = 'black',
                                            "grid_ensemble"="maroon", 
                                            "statn_ensemble"="darkturquoise", 
                                            "grid_all"="purple1",
                                            "statn_all"='springgreen4', 
                                            "all_ensemble" = "orange")) +
    ggtitle(calyear) +
    theme_bw()} %>%
    ggplotly()

@@

@@

```{r catch, echo = F}

setwd("C:/Users/sears/Documents/Repos/fourmile/catchments")

@@

## tmean 
tmean <- catchment %>%
  group_by(year, site) %>%
  summarize(tmean = mean(tmean)) %>%
  ggplot(aes(x = year, y=tmean, color = site)) +
  geom_point(size=3) +
  ggtitle('Average temp by year')

@@

## Niwot

### SWE

-Look at SWE dataframe as a group

@@

#look at swe -- ****** ask about this
niwot_swe <- niwot %>%
  dplyr::select(date, swe_mm_snotel, swe_mm_hrugrid, swe_mm_hrustatn) %>%
  mutate(swe_mm_snotel2 = swe_mm_snotel - lag(swe_mm_snotel))

@@

# niwot_swe <- niwot_swe %>% 
#   pivot_longer(!date, names_to = 'source', values_to = 'swe_mm')

@@

### ET

@@

>>>>

library(tidyverse)
library(plotly)
library(lubridate)
library(knitr)
library(kableExtra)
library(dataRetrieval)
library(hydroGOF)
library(sp)
library(raster)
library(sf)
library(mapview)
library(here)
library(scales)
library(dataRetrieval)

@@

```{r map, echo = F}

locations <- read_csv(here('data/GIS', 'fourmile_sensors.csv'))

hrus <- st_read(here('data/GIS', 'hrus_final.shp'))

locations <- locations %>%
  st_as_sf(., coords=c('y','x'), crs = 4326)

mapviewOptions(basemaps = c("Esri.WorldShadedRelief", "OpenStreetMap.DE"),
               layers.control.pos = "topright")

mapview(locations, zcol = 'type') + mapview(hrus)

@@

```{r boulder cr gage, echo = F}

#pull in Boulder Creek @ N 75th gage

bouldcr <- readNWISdv(siteNumbers = "06730200",
                          parameterCd = "00060", 
                          startDate = "2011-01-01", 
                          endDate = "2020-12-31")  %>%
  select(date = Date,
         q_cfs = X_00060_00003) %>%
  mutate(bouldcr = q_cfs/35.3146662) %>%
  select(-q_cfs) %>%
  mutate(year = year(date))

@@

q_ensemble <- ggplot(ensem_pivot, aes(date, q_m3s, color = cal_run)) +
  geom_line() +
  scale_color_manual(name="cal_run", values=c('obs'='black',
                                            "grid_ensemble"="maroon", 
                                            "statn_ensemble"="darkturquoise", 
                                            "grid_all"="purple1",
                                            "statn_all"='springgreen4', 
                                            "all_ensemble" = "orange")) +
  geom_line(data = bouldcr, aes(date, bouldcr/10), color = 'darkgrey') +
  scale_y_continuous(name = 'Q_m3s',
                    sec.axis = sec_axis(~.*10, name='BoulderCr_cms')) +
  theme_bw()

ggplotly(q_ensemble) %>%
  add_lines(data=bouldcr, x=~date, y=~bouldcr, colors=NULL, yaxis="y2", 
              inherit=FALSE, showlegend = FALSE) %>%
  layout(yaxis2 = list(overlaying = "y", side = "right",
                       color = 'darkgray',
                       title = "BoulderCr_cms"),
         legend = list(x = 1.1, y = 0.95))

@@

plotqensem_cy <- function(calyear) {
  {ensem_pivot %>%
    filter(year == calyear) %>%
    ggplot(aes(x = date,
               y = q_m3s,
               color = cal_run)) +
    geom_line() +
    scale_color_manual(name="cal_run", values=c('obs' = 'black',
                                            "grid_ensemble"="maroon", 
                                            "statn_ensemble"="darkturquoise", 
                                            "grid_all"="purple1",
                                            "statn_all"='springgreen4', 
                                            "all_ensemble" = "orange")) +
    ggtitle(calyear) +
    theme_bw()
    } %>%
    ggplotly()
}

@@

## Max Q prior to April 1

```{r}

ensem_bould <- left_join(ensem, bouldcr, by = 'date') %>%
  pivot_longer(!c(date, year), names_to = 'source', values_to = 'q_m3s') %>%
  mutate(month = month(date)) %>%
  filter(month %in% 1:3) %>%
  group_by(source, year) %>%
  slice_max(q_m3s, n=1)

early_peakq <- ensem_bould %>%
  mutate(doy = yday(date)) %>%
  select(c(doy, source, q_m3s, year)) %>%
  pivot_wider(names_from = 'source', values_from = c('q_m3s', 'doy'))

ggplot(early_peakq, aes(doy_bouldcr, doy_all_ensemble)) +
  geom_point(size=2.5) +
  geom_abline(intercept = 0, slope = 1, size=1, color='red') +
  theme_bw() +
  ggtitle('DOY of peak Q prior to April 1st - All Ensemble')

ggplot(early_peakq, aes(doy_bouldcr, doy_grid_all)) +
  geom_point(size=2.5) +
  geom_abline(intercept = 0, slope = 1, size=1, color='red') +
  theme_bw() +
  ggtitle('DOY of peak Q prior to April 1st - Grid All')

ggplot(early_peakq, aes(doy_bouldcr, doy_grid_ensemble)) +
  geom_point(size=2.5) +
  geom_abline(intercept = 0, slope = 1, size=1, color='red') +
  theme_bw() +
  ggtitle('DOY of peak Q prior to April 1st - Grid Ensemble')

ggplot(early_peakq, aes(doy_bouldcr, doy_statn_all)) +
  geom_point(size=2.5) +
  geom_abline(intercept = 0, slope = 1, size=1, color='red') +
  theme_bw() +
  ggtitle('DOY of peak Q prior to April 1st - Station All')

ggplot(early_peakq, aes(doy_bouldcr, doy_statn_ensemble)) +
  geom_point(size=2.5) +
  geom_abline(intercept = 0, slope = 1, size=1, color='red') +
  theme_bw() +
  ggtitle('DOY of peak Q prior to April 1st - Station Ensemble')

early_peakq_table <- ensem_bould %>%
  mutate(doy = yday(date)) %>%
  dplyr::arrange(date)

kable(early_peakq_table) %>%
  kable_styling(full_width = F) %>% 
  scroll_box(width = "100%", height = "500px")

@@

## Precip

```{r catch precip, echo = F}
setwd("C:/Users/sears/Documents/Repos/fourmile/catchment_output")

@@

## Tmean

```{r tmean catch, echo = F}
## tmean 
tmean <- catchment %>%
  group_by(year, site) %>%
  summarize(tmean = mean(tmean)) %>%
  ggplot(aes(x = year, y=tmean, color = site)) +
  geom_point(size=3) +
  ggtitle('Average temp by year')

@@

# Niwot and HRU7

## SWE and precip

@@

# look at swe
niwot_swe <- niwot %>%
  dplyr::select(date, swe_mm_snotel, swe_mm_hrugrid, swe_mm_hrustatn)

@@

# cumulative swe
niwot_swe2 <- niwot_swe %>%
  mutate(year = year(date)) %>%
  pivot_longer(!c(date,year), names_to = 'source', values_to = 'swe_mm') %>%
  group_by(year, source) %>%
  mutate(swe_cumsum_mm = cumsum(swe_mm))

cum_swe_niwot <- ggplot(niwot_swe2, aes(date, swe_cumsum_mm, color = source)) +
  geom_line() +
  scale_x_date(labels = date_format("%b")) +
  facet_wrap(~year, scales = 'free')

ggplotly(cum_swe_niwot)

#look at precip
niwot_p <- niwot %>%
  dplyr::select(date, precip_mm_snotel, precip_mm_hrugrid, precip_mm_hrustatn)

kable(niwot_p) %>%
  kable_styling(full_width = F) %>% 
  scroll_box(width = "100%", height = "500px")

# cumulative precip
niwot_p2 <- niwot_p %>%
  mutate(year = year(date)) %>%
  pivot_longer(!c(date,year), names_to = 'source', values_to = 'precip_mm') %>%
  group_by(year, source) %>%
  mutate(p_cumsum_mm = cumsum(precip_mm))

cum_precip_niwot <- ggplot(niwot_p2, aes(date, p_cumsum_mm, color = source)) +
  geom_line() +
  scale_x_date(labels = date_format("%b")) +
  facet_wrap(~year, scales = 'free')

ggplotly(cum_precip_niwot)

@@

## ET

@@

# 1:1 scatter plots for Niwot vs grid and Niwot vs station
niwot_et_scatter <- niwot_et_compare %>%
  filter(et_mm > 0)

# Niwot vs grid
et_grid_obs <- niwot_et_scatter %>%
  filter(source == c('et_niwot','et_hru7grid')) %>%
  pivot_wider(names_from = 'source', values_from = 'et_mm')

#scatter of grid vs niwot
et_scatter_grid_obs <- ggplot(et_grid_obs, aes(x = et_niwot, y = et_hru7grid)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, size=1, color='red') + 
  ggtitle('Niwot ET vs. HRU7 grid ET')

et_scatter_grid_obs

# Niwot vs statn inputs
et_statn_obs <- niwot_et_scatter %>%
  filter(source == c('et_niwot','et_hru7statn')) %>%
  pivot_wider(names_from = 'source', values_from = 'et_mm')

#scatter of grid vs niwot
et_scatter_statn_obs <- ggplot(et_statn_obs, aes(x = et_niwot, y = et_hru7statn)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, size=1, color='red') + 
  ggtitle('Niwot ET vs. HRU7 station ET')

et_scatter_statn_obs
